---
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2: 
    template: templates/brief_template.tex
documentclass: book
bibliography: references.bib
---
  
# Customisations and extensions

This chapter describes a number of possible customizations to the `oxforddown` thesis. 

## Embedding PDF documents as chapters

You may want to embed existing PDF documents into the thesis, for example if your department allows a 'portfolio' style thesis and you need to include an existing typeset publication as a chapter. 

In gitbook output, you can simply use `knitr::include_graphics` and it should include a scrollable (and downloadable) PDF.
You will probably want to set the chunk options `out.width='100%'` and `out.height='1000px'`:

```{r, eval=knitr::is_html_output(), out.width='100%', out.height='1000px'}
knitr::include_graphics("figures/pdf_example/Lyngs2020_FB.pdf")
```
<br>

In LaTeX output, however, this approach can cause odd behaviour.
Therefore, when you build your thesis to PDF, split the PDF into an alphanumerically sorted sequence of **single-page** PDF files (you can do this automatically with the package `pdftools`). You can then use the appropriate LaTeX command to insert them, as shown below (for brevity, in the `oxforddown` PDF sample content we're only including two pages)
*Note that the chunk option `results='asis'` must be set.*
You may also want to remove margins from the PDF files, which you can do with Adobe Acrobat (paid version) and likely other software.

```{r, echo=TRUE, results='asis', eval=knitr::is_latex_output(), linewidth = 70}
# install.packages(pdftools)
# split PDF into pages stored in figures/pdf_example/split/
# pdftools::pdf_split("figures/pdf_example/Lyngs2020_FB.pdf",
#        output = "figures/pdf_example/split/")

# grab the pages
pages <- list.files("figures/pdf_example/split", full.names = TRUE)

# set how wide you want the inserted PDFs to be: 
# 1.0 is 100 per cent of the oxforddown PDF page width;
# you may want to make it a bit bigger
pdf_width <- 1.2

# for each PDF page, insert it nicely and
# end with a page break
cat(stringr::str_c("\\newpage \\begin{center} \\makebox[\\linewidth][c]{\\includegraphics[width=", pdf_width, "\\linewidth]{", pages, "}} \\end{center}"))

```

## Customizing referencing 

### Using a .csl file with pandoc instead of biblatex
The `oxforddown` package uses biblatex in latex for referencing. 
It is also possible to use pandoc for referencing by providing a .csl file in the YAML header of **index.Rmd** (likely requiring commenting out the biblatex code in **templates/template.tex**). 
This may be helpful for those who have a .csl file describing the referencing format for a particular journal. 
However, note that this approach does not support chapter bibliographies (see Section \@ref(biblatex-custom)).

```YAML
csl: ecology.csl
```

### Customizing biblatex and adding chapter bibliographies {#biblatex-custom}

This section provides one example of customizing biblatex. Much of this code was combined from searches on Stack Exchange and other sources (e.g. [here](https://tex.stackexchange.com/questions/10682/suppress-in-biblatex)).

In **templates/template.tex**, one can replace the existing biblatex calls with the following to achieve referencing that looks like this: 

(Charmantier and Gienapp 2014)

Charmantier, A. and P. Gienapp (2014). Climate change and timing of avian breeding and migration: evolutionary versus plastic changes. Evolutionary Applications 7(1):15–28. doi: 10.1111/eva.12126.


```latex
\usepackage[backend=biber,
    bibencoding=utf8,
    refsection=chapter, % referencing by chapter
    style=authoryear, 
    firstinits=true,
    isbn=false,
    doi=true,
    url=false,
    eprint=false,
    related=false,
    dashed=false,
    clearlang=true,
    maxcitenames=2,
    mincitenames=1,
    maxbibnames=10,
    abbreviate=false,
    minbibnames=3,
    uniquelist=minyear,
    sortcites=true,
    date=year
]{biblatex}
\AtEveryBibitem{%
  \clearlist{language}%
  \clearfield{note}
}

\DeclareFieldFormat{titlecase}{\MakeTitleCase{#1}}

\newrobustcmd{\MakeTitleCase}[1]{%
  \ifthenelse{\ifcurrentfield{booktitle}\OR\ifcurrentfield{booksubtitle}%
    \OR\ifcurrentfield{maintitle}\OR\ifcurrentfield{mainsubtitle}%
    \OR\ifcurrentfield{journaltitle}\OR\ifcurrentfield{journalsubtitle}%
    \OR\ifcurrentfield{issuetitle}\OR\ifcurrentfield{issuesubtitle}%
    \OR\ifentrytype{book}\OR\ifentrytype{mvbook}\OR\ifentrytype{bookinbook}%
    \OR\ifentrytype{booklet}\OR\ifentrytype{suppbook}%
    \OR\ifentrytype{collection}\OR\ifentrytype{mvcollection}%
    \OR\ifentrytype{suppcollection}\OR\ifentrytype{manual}%
    \OR\ifentrytype{periodical}\OR\ifentrytype{suppperiodical}%
    \OR\ifentrytype{proceedings}\OR\ifentrytype{mvproceedings}%
    \OR\ifentrytype{reference}\OR\ifentrytype{mvreference}%
    \OR\ifentrytype{report}\OR\ifentrytype{thesis}}
    {#1}
    {\MakeSentenceCase{#1}}}
    
% \renewbibmacro{in:}{}
% suppress "in" for articles
% 
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
%-- no "quotes" around titles of chapters/article titles
\DeclareFieldFormat[article, inbook, incollection, inproceedings, misc, thesis, unpublished]
{title}{#1}
%-- no punctuation after volume
\DeclareFieldFormat[article]
{volume}{{#1}}
%-- puts number/issue between brackets
\DeclareFieldFormat[article, inbook, incollection, inproceedings, misc, thesis, unpublished]
{number}{\mkbibparens{#1}} 
%-- and then for articles directly the pages w/o any "pages" or "pp." 
\DeclareFieldFormat[article]
{pages}{#1}
%-- for some types replace "pages" by "p."
\DeclareFieldFormat[inproceedings, incollection, inbook]
{pages}{p. #1}
%-- format 16(4):224--225 for articles
\renewbibmacro*{volume+number+eid}{
  \printfield{volume}%
  \printfield{number}%
  \printunit{\addcolon}
}
```

If you would like chapter bibliographies, in addition insert the following code at the end of each chapter, and comment out the entire REFERENCES section at the end of template.tex.

```latex
\printbibliography[segment=\therefsection,heading=subbibliography]
```

## Customizing the page headers and footers

The following code, when it replaces the existing correpsonding code block in **ociamthesis.cls**, puts chapter number and title centered in the header and page number in the footer, centered. 
This may be desirable particularly when inserting PDF chapters, as the margins of the PDF may not exactly align with the left and right margins of the page, demarcated by the existing header and footer text. 
In the following code block, the original code is commented out where replaced.

```latex
\usepackage{fancyhdr}
\setlength{\headheight}{15pt}
\fancyhf{} % clear the header and footers
\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{\markboth{\thechapter. #1}{\thechapter. #1}}
% \renewcommand{\chaptermark}[1]{\markboth{\thechapter. #1}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection. #1}}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[CO]{\emph{\leftmark}}
\fancyhead[CE]{\emph{\rightmark}}
% \fancyhead[LO,RE]{}
% \fancyhead[LE,RO]{}
\fancyfoot[CO,CE]{\emph{\thepage}}

\fancypagestyle{plain}{\fancyhf{}\fancyfoot[C]{\emph{\thepage}}}

% JEM fix header on cleared pages for openright
\def\cleardoublepage{\clearpage\if@twoside \ifodd\c@page\else
   \hbox{}
   % \fancyhead[RE,LO]{}
   \fancyhead[CE,CO]{}
   \newpage
   \if@twocolumn\hbox{}\newpage
   \fi
   % \fancyhead[LO]{\emph{\leftmark}} 
   % \fancyhead[RE]{\emph{\rightmark}} 
   \fancyhead[CO]{\emph{\leftmark}} 
   \fancyhead[CE]{\emph{\rightmark}} 
   \fi\fi}
```
